Statement of Intent

### 2010-12-15 applications/phd_programs/personal_statement.tex
Done with a generic personal statement

\documentclass[onecolumn, 12pt]{article}
\usepackage{fullpage}
\usepackage{graphicx}

\title{}
\author{}
\date{}

@@

In my second year of college I started as a research assistant in the Applied
Machine Learning lab at Brigham Young University.  The project that was given
to me to work on was particle swarm optimization (PSO).  While my work in PSO
did not relate directly to my interest in machine interpretation, it taught me
what it is like to do real research, as well as the joys and sorrows of the
paper submission process at conferences and journals.

All told I spent two and a half years working on various aspects of PSO.  At
the beginning I was just running experiments with code other people wrote, a
simple bookkeeper.  I progressed to helping analyze and interpret the results,
then to writing a little code of my own.  After about a year in the lab, I got
my first publication; I was a co-author on a paper about how parallelization
affects how PSO~\cite{mcnabb-2009-large-particle-swarms}.

@@

I had my first major setback when that paper was rejected outright.  At this
point I had spent two years working on PSO, and I felt like giving up and
moving on.  PSO wasn't my main interest anyway, and I wanted to do more
interesting things.  My advisor, however, insisted that I persevere, telling
me that the experience would be good practice for a PhD.  So I revised my paper
and submitted it to another top conference, where it was
accepted~\cite{gardner-2010-speculative-evaluation-in-pso}.

@@

I have had many other experiences that also contribute to my preparation for
your program.  I have gained a broad background in statistical machine learning
techniques, having been the TA for the Bayesian Inference class taught at BYU
and taken several advanced classes on the subject.  I have also taken all of
the classes that BYU offers in natural language processing (NLP).  For my final
project in the graduate-level NLP course, I implemented a Gibbs sampler for the
latent Dirichlet allocation model and ran it on a novel dataset of religious
discourses.  That work was of particular interest to BYU, and I presented it at
the BYU Studies Symposium, an international conference held at
BYU~\cite{gardner-2010-general-conference-topics}.

As a masters student at BYU I have continued my work with topic modeling,
exploring how topic models can be used to improve the browsing experience of
large corpora.  I recently returned from the NIPS conference, where I presented
at a workshop a tool that I and others at BYU have built that incorporates a
lot of previous work in visualizing topic models and some of our own
ideas~\cite{gardner-2010-topic-browser}.

@@

\bibliographystyle{unsrt}
\bibliography{bib}

>>>>

\documentclass[onecolumn, 12pt]{article}
\usepackage{fullpage}
\usepackage{graphicx}
\usepackage{natbib}

\setlength{\bibsep}{0in}

@@

In my second year of college I started as a research assistant in the Applied
Machine Learning lab at Brigham Young University (BYU).  The project that was
given to me to work on was particle swarm optimization (PSO).  While my work in
PSO did not relate directly to my interest in machine interpretation, it taught
me what it is like to do real research and prepared me for my future work as a
PhD student.

All told I spent two and a half years working on various aspects of PSO.  At
the beginning I was just running experiments with code other people wrote, a
simple bookkeeper.  I progressed to helping analyze and interpret the results,
then to writing a little code of my own.  After about a year in the lab, I got
my first publication; I was a co-author on a paper about parallelization and
scalability in PSO~\cite{mcnabb-2009-large-particle-swarms}.

@@

I had my first major setback when that paper was rejected outright.  At this
point I had spent two years working on PSO, and I felt like giving up and
moving on.  My advisor, however, insisted that I persevere, telling me that the
experience would be good practice for a PhD.  So I revised my paper and
submitted it to another top conference, where it was
accepted~\cite{gardner-2010-speculative-evaluation-in-pso}.

@@

I have had many other experiences that also contribute to my preparation for
your program.  I have gained a broad background in statistical machine learning
techniques, having been the TA for the Bayesian Inference class and taken
several advanced classes on the subject.  I have also taken all of the classes
that BYU offers in natural language processing (NLP).  For my final project in
the graduate-level NLP course, I implemented a Gibbs sampler for the latent
Dirichlet allocation model and ran it on a novel dataset of religious
discourses.  That work was of particular interest to BYU, and I presented it at
the BYU Studies Symposium, an international conference held at the
university~\cite{gardner-2010-general-conference-topics}.

At the end of my undergraduate education I applied for and received a
scholarship from the North American chapter of the Association for
Computational Linguistics (NAACL) to attend a summer school at Johns Hopkins
University on human language technologies.  My experience at JHU over the
summer was further preparation for your program, as we covered a broad range of
topics dealing with machine learning and natural language processing.  I also
attended the NAACL conference that summer and saw firsthand some of the
research going on at your school.

As a masters student at BYU I have continued my work with topic modeling,
exploring how topic models can be used to improve the browsing experience of
large corpora.  I recently returned from the NIPS conference, where I presented
a tool that I and others at BYU have built that incorporates a lot of previous
work in visualizing topic models and some of our own
ideas~\cite{gardner-2010-topic-browser}.

@@

\vspace{2mm}
\footnotesize
\bibliographystyle{plain}
\renewcommand\bibsection{\noindent \small\textbf{Publications}\vspace{-2mm}\footnotesize}
\bibliography{bib}
